Turbo fast no compromise gravitational wave parameter estimation

### 2023-01-06 src/tex/ms.tex
Update algorithm part
Address some of Tom's comments

One may be concerned by the accuracy of this approximation over the target
parameter space, especially in the region where the generated waveform is
significantly different from the reference waveform. However, given we are
interested in the most probable set of parameters, if we choose the reference
waveform to be close to the data, the waveforms that are different from the
reference waveform should also differ significantly from the data. This means
the likelihood value for these waveforms should be significantly smaller than
the likelihood of the waveforms that are similar to the reference waveform,
hence will not be relevant for the PE result. In practice, one will first
optimize the likelihood function with full frequency resolution to obtain the
reference waveform parameters, which can be run at a much lower cost compared to
PE. \te{I find this point a little unclear. I don't think there is any guarentee that the likelihood values far from the reference waveform will not be overestimated using the method right? Do you therefore have to explicitly ensure you're sampler doesn't wander too far from the reference waveform?}

@@

The final ingredient we need is the heterodyning coefficients given the data and
the reference waveform on the sparse bins, which are given by
\begin{align}
    A_0(b) &= 4 \sum_{f \in b} \frac{d(f)h^*_0(f)}{S_n(f)} df, \\
    A_1(b) &= 4 \sum_{f \in b} \frac{d(f)h^*_0(f)(f-f_m(b))}{S_n(f)} df, \\
    B_0(b) &= 4 \sum_{f \in b} \frac{|h_0(f)|^2}{S_n(f)} df, \\
    B_1(b) &= 4 \sum_{f \in b} \frac{|h_0(f)|^2(f-f_m(b))}{S_n(f)} df.
\end{align}
Note that the sum within each bin should be done with the same sampling rate as
the data, i.e. the same one would do in the case without using heterodyned
likelihood. \te{If we want this to be maximally instructive, it might be useful to include a plot here. Something like the Fig 1 from the Zackay paper but maybe explicitly showing the two parts of the integral i.e. the highly oscillating part and the smooth part?}

@@

% The flowMC algorithm
As mentioned in the previous subsection, one of the main issue gradient-based
sampler is it does not explore locally correlated features and multi-modality well. This
is exactly where the normalizing flow model can help. Once multiple independent
chains have generated samples, we can combine the samples and feed them to the
normalizing flow, which learns the global landscape of the target
distribution. Note that since the independent chains are generated with the same
number of samples, the normalization between posterior density represented by
different chains is in general different from the true target distribution. To
adjust for this inaccuracy, we use the normalizing flow model as a
\textit{proposal} distribution. Given a trained normalizing flow model, we can
generate the proposed jump in the target space by sampling from the latent
distribution $z \sim p_z(Z)$, then push it through the learned map given by the
normalizing flow model $x=f(z)$. the acceptance criteria is given by
\begin{align}
    \alpha(\mathbf{x},\mathbf{x_0}) = \min \left[ 1, \frac{\hat{\rho}(\mathbf{x_0})\rho_*(\mathbf{x})}{\hat{\rho}(\mathbf{x})\rho_*(\mathbf{x_0})}\right],
\end{align}
where $\hat{\rho}$ is the probability density estimated by the normalizing flow
model, $\rho_*$ is the probability density evaluated using the target function,
and $x_0$ is the current position. We can see the flow distribution is the
target distribution when the accepting probability is 1. When the normalizing
flow model has not converged to the target distribution, only a portion of the
proposed jumps will be accepted. This means an MCMC process using the normalizing
flow model as the proposal distribution can adjust the normalization across
different regions of the target parameter space by rejecting jumps into less
likely regions. The training and sampling are then repeated until certain
criteria are met. Note that every time we retrain the network, we are breaking the
Markov properties since we are changing the proposal distribution. To produce
final samples that can be used to estimate target quantities, one has to freeze
the normalizing flow model and not retrain during the final sampling phase.
To summarize the sampling process, the pseudocode is given by

\begin{algorithm}
\caption{flowMC pseudocode}\label{alg:cap}
\KwIn{Initial position $ip$}
\Parameters{n training loop $nt$, n production loop $np$}
\Variables{current chain $cc$, current position $cp$, current NF parameters $\Theta$}
\KwResult{Chains}
$cp \leftarrow ip$\\
\tcc{Training loop}
    \For{$i<nt$}{
        $cc, cp \leftarrow LocalSampling(cp)$\\
        $\Theta \leftarrow TuneNF(cc)$
        $chains, cp \leftarrow GlobalSampling(cp, \Theta)$ 
        $cc \leftarrow Append(cc, chains)$
    }
\tcc{Production loop}
    \For{$i<np$}{
        $Chains, cp \leftarrow LocalSampling(cp)$
        $subchains, cp \leftarrow GlobalSampling(cp, \Theta)$ 
        $Chains \leftarrow Append(cc, chains)$
    }

@@

\te{Discussion of detailed balance here? I know that some people really care about this stuff}
\subsection{Accelerators}
\label{sec:accelerators}

@@

\section{Acknowledgements}

>>>>

One may be concerned by the accuracy of this approximation over the target
parameter space, especially in the region where the generated waveform is
significantly different from the reference waveform. However, given we are
interested in the most probable set of parameters, if we choose the reference
waveform to be close to the data, the waveforms that are different from the
reference waveform should also differ significantly from the data. This means
the likelihood value for these waveforms should be significantly smaller than
the likelihood of the waveforms that are similar to the reference waveform,
hence will not be relevant for the PE result. In practice, one will first
optimize the likelihood function with full frequency resolution to obtain the
reference waveform parameters, which can be run at a much lower cost compared to
PE.

@@

The final ingredient we need is the heterodyning coefficients given the data and
the reference waveform on the sparse bins, which are given by
\begin{align}
    A_0(b) &= 4 \sum_{f \in b} \frac{d(f)h^*_0(f)}{S_n(f)} df, \\
    A_1(b) &= 4 \sum_{f \in b} \frac{d(f)h^*_0(f)(f-f_m(b))}{S_n(f)} df, \\
    B_0(b) &= 4 \sum_{f \in b} \frac{|h_0(f)|^2}{S_n(f)} df, \\
    B_1(b) &= 4 \sum_{f \in b} \frac{|h_0(f)|^2(f-f_m(b))}{S_n(f)} df.
\end{align}
Note that the sum within each bin should be done with the same sampling rate as
the data, i.e. the same one would do in the case without using heterodyned
likelihood.

@@

% The flowMC algorithm
As mentioned in the previous subsection, one of the main issue gradient-based
sampler is it does not explore locally correlated features and multi-modality well. This
is exactly where the normalizing flow model can help. Once multiple independent
chains have generated samples, we can combine the samples and feed them to the
normalizing flow, which learns the global landscape of the target
distribution. Note that since the independent chains are generated with the same
number of samples, the normalization between posterior density represented by
different chains is in general different from the true target distribution. To
adjust for this inaccuracy, we use the normalizing flow model as a
\textit{proposal} distribution. Given a trained normalizing flow model, we can
generate the proposed jump in the target space by sampling from the latent
distribution $z \sim p_z(Z)$, then push it through the learned map given by the
normalizing flow model $x=f(z)$. the acceptance criteria is given by
\begin{align}
    \alpha(\mathbf{x},\mathbf{x_0}) = \min \left[ 1, \frac{\hat{\rho}(\mathbf{x_0})\rho_*(\mathbf{x})}{\hat{\rho}(\mathbf{x})\rho_*(\mathbf{x_0})}\right],
\end{align}
where $\hat{\rho}$ is the probability density estimated by the normalizing flow
model, $\rho_*$ is the probability density evaluated using the target function,
and $x_0$ is the current position. We can see the flow distribution is the
target distribution when the accepting probability is 1. When the normalizing
flow model has not converged to the target distribution, only a portion of the
proposed jumps will be accepted. This means an MCMC process using the normalizing
flow model as the proposal distribution can adjust the normalization across
different regions of the target parameter space by rejecting jumps into less
likely regions. The training and sampling are then repeated until certain
criteria are met. Note that every time we retrain the network, we are breaking the
Markov properties since we are changing the proposal distribution. To produce
final samples that can be used to estimate target quantities, one has to freeze
the normalizing flow model and not retrain during the final sampling phase in order to satisfy the detailed balance condition.
To summarize the sampling process, the pseudocode is given by

\begin{algorithm}
\caption{flowMC pseudocode}\label{alg:cap}
\KwIn{Initial position $ip$}
\Parameters{n training loop $nt$, n production loop $np$}
\Variables{current chain $cc$, current position $cp$, current NF parameters $\Theta$}
\KwResult{Chains}
$cp \leftarrow ip$\\
\tcc{Training loop}
    \For{$i<nt$}{
        $cc, cp \leftarrow LocalSampling(cp)$\\
        $\Theta \leftarrow TuneNF(cc)$\\
        $chains, cp \leftarrow GlobalSampling(cp, \Theta)$ \\
        $cc \leftarrow Append(cc, chains)$
    }
\tcc{Production loop}
    \For{$i<np$}{
        $c_local, cp \leftarrow LocalSampling(cp)$\\
        $c_global, cp \leftarrow GlobalSampling(cp, \Theta)$ \\
        $Chains \leftarrow Append(Chains, c_local, c_global)$
    }

@@

\subsection{Accelerators}
\label{sec:accelerators}

@@

\section{Acknowledgements}
Thanks Will.
